---
title: "Things to Know for Week 7 (Chapter 6)"
output: 
  html_document:
    theme: cerulean
    highlight: pygments
    css: lab.css
---

```{r setup, include = FALSE}
library(tidyverse)
library(tidytext)
library(gutenbergr)
library(janeaustenr)
library(scales)
library(wordcloud)
library(reshape2)
```

Here are a few reminders and some things to be on the look-out for in Chapter 6.

  + **NEW PACKAGES:** The following packages are new this week and will need to be installed prior to use: `topicmodels`
  + **GREAT LIBRARY HEIST:** I had some trouble with the *Pride and Prejudice* text. This book, from Project Gutenberg includes a table of contents, and leading whitespaces in the text column. Here's how I got around the issue.  
    ```{r, echo = TRUE, eval = FALSE}
    #Read in the thre other texts
    titles <- c("Twenty Thousand Leagues under the Sea",
            "The War of the Worlds",
            "Great Expectations")

    books <- gutenberg_works(title %in% titles) %>%
      gutenberg_download(meta_fields = "title", mirror = "http://mirrors.xmission.com/gutenberg/")
    
    #Read in Pride and Prejudice separately
    pride <- gutenberg_download(gutenberg_id = 1342)
    
    #Add in the title column -- rep() stands for repeat, so we are just making
    #a column of the appropriate length here.
    pride$title <- rep("Pride and Prejudice", nrow(pride))
    
    #Cut out the rows corresponding to the table of contents.
    pride <- pride[c(1:12, 139:nrow(pride)), ]
    
    #Remove the leading whitespace in the text column -- we'll *apply* the
    #trimws function to every cell in the text column of our
    #pride data frame. The result of lapply() is a list, so we will just
    #unlist() the result to get back to a simple string value. (There are
    #probably beter ways to do this).
    pride$text <- unlist(lapply(pride$text, trimws))
    
    #Combine the original three books with the Pride and Prejudice text.
    books <- bind_rows(books, pride)
    ```
  
***

That's all for now. Post questions you have in Slack and/or bring them with you to our Thursday live meeting.

***

<p style = "text-align:left;">[Previous, Read Chapter Five](https://www.tidytextmining.com/dtm.html)
<span style = "float:right;">[Next, Read Chapter Six](https://www.tidytextmining.com/topicmodeling.html)</span>
</p>